---
title: 'Lab Week 5: time series and A MAP!'
author: "Amelia Ritger"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

##3 Attach packages
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

### Monthly US energy consumption (renewables)
```{r}
us_renew <- read_csv(here::here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

unique(us_renew$description)
```

- Make the description all lowercase
- Only keep observations where "description" variable contains "consumption"
- Remove any observations where "description" vraiable contains "total"

```{r}
renew_clean <- us_renew %>%
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern="consumption")) %>% 
  filter(!str_detect(description, pattern="total"))
```

### Convert `yyyymm` to a date
```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = tsibble::yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# make a version where I have the month and year in separate columns
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

### Look at it:
```{r}
renew_gg <- ggplot(data=renew_date, aes(x=month_sep, y=value)) +
  geom_line(aes(color=description))

renew_gg
```

Updating colors with paleteer palettes:
```{r}
view(palettes_d_names)

renew_gg + 
  scale_color_paletteer_d("nationalparkcolors::CraterLake")
```

### Coerce renew_parsed to a tsibble
```{r}
renew_ts <- as_tsibble(renew_parsed, key=description, index= month_sep)
```

Let's look at our time series data in a few different ways:
```{r}
#these only work if your df is a tsibble
renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
renew_ts %>% gg_season(value)
#Let's make this graph with ggplot!

ggplot(data=renew_parsed, aes(x=month, y=value, group=year)) +
  geom_line(aes(color=year)) +
  facet_wrap(~description,
             ncol=1,
             scales="free",
             strip.position = "right")
```

### Just look at the hydroelectric energy consumption
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)
#let's make this again in ggplot

ggplot(hydro_ts, aes(x=month, y=value, group=year)) +
  geom_line(aes(color=year))
```

### What if I want the quarterly average consumption for hydro?
```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>%  #period means based off different groups that exist
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)
```

### Decompose that hydro_ts
```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window=5))) #window here is essentially the lag for the seasonal component (hence not homogenous across the board)

components(dcmp) %>% autoplot()
```

```{r}
hist(components(dcmp)$remainder)
```

Now let's look at the ACF:
```{r}
hydro_ts %>% 
  feasts::ACF(value) %>% 
  autoplot()
```

# **DANGER DANGER**

### read a lot more before throwing this function (to make forecasts) around
```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h="4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep)>2010))
```

### Let's make a world map!
```{r}
world <- read_sf(dsn=here("data","TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer="TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```



